---
title: "Covid91 vaccine study Final2023"
author: "Seth Guzman"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

data from this study is taken from FinalData


$H_0$: the vaccines has no effect on people within a subgroup

$H_A$: the vaccine has an effect on people within a subgroup



## Methods

The statistical analysis i will be using to sort the data will be cat~cat
because we are putting the sub groups into category 



the methods we will use to display the data is  bar chart, cross table, and chi squared test

## sub categorize male data
```{r}
maleData = filter (FinalData,sex == "M")
```


## Males


## Graphical Results
```{r} 
barchartGC(~infected+treatment,data=maleData)
barchartGC(~infected+treatment,data=maleData, type="percent")

```
This graph shows use that from the all male group of people who contracted the ones who received a vaccine had less symptoms then those with the placebo while the all male group that didnt contract covid showed no difference between those who were vaccinated and un vaccinated.
## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=maleData)
rowPerc(table1)
colPerc(table1)
```

The data shows us that 69.57% of the unvaccinated male group experienced symptons while only 30.43% of the vaccinated males experienced symptoms

It also tells us that from the male placebo group 2.55% of them contracted covid while only 1.15% of the vaccinated group contracted covid

## Inferential Result


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p value is 4.631e-14  making it to small to be in significant therfore disproving the null hypothesis

Based on Fishers exact value You become .35% more susceptible to contracting covid if you are unvaccinated 





## sub categorizing female data

```{r}
femaleData = filter (FinalData,sex == "F")
```


## Females

## Graphical Results

```{r}
barchartGC(~infected + treatment, data=femaleData)
barchartGC(~infected+treatment, data=femaleData, type="percent")

```

This graph shows use that from the all female group of people who contracted the ones who received a vaccine had less symptoms then those with the placebo while the all female group that didnt contract covid showed no difference between those who were vaccinated and un vaccinated.


## Numerical Results

```{r}
table2 <- xtabs(~infected + treatment, data=femaleData)
rowPerc(table2)
colPerc(table2)
```
The data from this table shows that from the female group that were infected with covid-19 of those who were unvaxed 60% of them experienced symptoms while only 40% of people in the vaccinated group experienced symptoms 

This data also depicts that within the female placebo group 2.04% were infected with covid but only 1.37% of the vaccinated group 




## Inferential Result

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```
The p value within the female group is  0.0002162 showing that theres a insignificant differences so i am failing to reject the null hyothesis and conclude the covid vaccine is inneftive within the female population

## subsetting lgtbq data

```{r}
LGBTQData = filter (FinalData,LGBTQ == "gay")
```
## LGBTQ
## Graphical Results

```{r}
barchartGC(~infected+treatment, data=LGBTQData)
barchartGC(~infected+treatment, data=LGBTQData, type="percent")

```

This Graph shows that within the lgbtq subgroup Those afflicted with covid 19 and that were vaccinated experienced more symptoms then their un vaccinated counter part. for the people who never contracted covid vaxxed or un-vaxxed it didn't show any difference 


## Numerical Results

```{r}
table3 <- xtabs(~infected + treatment, data=LGBTQData)
rowPerc(table3)
colPerc(table3)
```

This table shows that within the all gay group those who werent vaccinated had a 15.38% infection rate while those who were vaccinated had a 84.62% infection rate. 

From the all gay placebo group only .66% of the unvaccinated got infected with covid-19 while a large 3.54% of vaccinated gay people got infected with covid_19


## Inferential Result
```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

The p value within the LGBTQ group is  0.0000006668  showing that theres a significant difference so i am rejecting the null hyothesis and conclude the covid vaccine has an effect within the LGBTQ population


## subset druggie data
```{r}
DrugUserData = filter(FinalData, DrugUser == "yes")
```
## Druggies
## Graphical Results

```{r}
barchartGC(~infected + treatment, data=DrugUserData)
barchartGC(~infected+treatment, data=DrugUserData, type="percent")

```

This graph shows that for druggies who were infected wih covid-19 those who were vaccinated actually experienced more symptoms then those that were unvaccinated. The druggies who didnt get infected with covid-19 wether vaccinated or unvaccinated showed no difference


## Numerical Results

```{r}
table4 <- xtabs(~infected + treatment, data=LGBTQData)
rowPerc(table4)
colPerc(table4)
```

## Inferential Result

```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```

# Overall Results and Conclusions
